Efficient and Adjustable Run-Time Monitoring Using Dataflow-Guided Filtering

### 2014-09-12 evaluation.tex
Added discussion on random dropping to policies.

We tested our system using all C benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 2 billion instructions. An initial slack of 2
million cycles, which is less than 1\% of
the total execution time, is given for a 10\% target overhead. This initial
slack is scaled proportionally for other overhead targets.

@@

% UMC coverage across policies
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage for uninitialized memory check for different dropping policies.}
    \label{fig:evaluation.umc_coverage}
    \vspace{-0.1in}
  \end{center}
\end{figure}
In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of independent monitoring flows.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that source-dropping
consistently achieves higher coverage than unrestricted dropping. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.events}.

@@

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by randomly deciding whether events
should be dropped.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running for
unrestricted and source-only dropping policies. Here, we use a 10\% probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored. Each run
was simulated for 200 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that since there are multiple dependent
monitoring operations that must be done in order for a final check to be
successful, a small amount of randomness at each event quickly compounds to
reduce the probability of certain checks to near zero. Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 5\% increase in coverage with only seven runs.
Over hundreds or thousands of runs, source-only dropping should be able to
reach 100\% code coverage.

>>>>

We tested our system using all C benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 2 billion instructions. An initial slack of 2
million cycles, which is less than 1\% of
the total execution time, is given for a 10\% target overhead. This initial
slack is scaled proportionally for other overhead targets. Unless otherwise
specified, an unrestricted, slack-based dropping policy is used.

@@

% UMC coverage across policies
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage for uninitialized memory check for different dropping policies.}
    \label{fig:evaluation.umc_coverage}
    \vspace{-0.1in}
  \end{center}
\end{figure}
In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of independent monitoring flows.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that source-dropping
consistently achieves higher coverage than unrestricted dropping. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

@@

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running for
unrestricted and source-only dropping policies with probabilistic dropping. Here, we use a 10\% probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored. Each run
was simulated for 200 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that since there are multiple dependent
monitoring operations that must be done in order for a final check to be
successful as discussed in Section~\ref{sec:policies.when}.
Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 5\% increase in coverage with only seven runs.
Over hundreds or thousands of runs, source-only dropping should be able to
reach 100\% code coverage.

### 2014-09-12 policies.tex
Added discussion on random dropping to policies.

Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
made. Specifically, we first discuss how to measure the run-time monitoring
overheads and when monitoring events need to be dropped in
Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
selecting which events are dropped (Section~\ref{sec:policies.events}). Different policies in determining which
events can be dropped create a trade-off between how closely the overhead
budget is met and the monitoring coverage achieved.

\subsection{Measuring Monitoring Overheads}
\label{sec:policies.slack}

@@

% Slack tracking module
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/stm.pdf}
    \vspace{-0.2in}
    \caption{Slack tracking and drop decision hardware.}
    \label{fig:policies.stm}
    \vspace{-0.1in}
  \end{center}
\end{figure}

In order to decide whether a monitoring event should be dropped or not, we need
a way to keep track of overheads. In this section, we present a specific method
for measuring execution time overheads. However, we note that other methods
could be used in order to target budgets on IPC, power, energy, etc.  
% For example, one way to measure execution time overheads would be to insert
% checkpoints in the program. By recording the execution time of these
% checkpoints without monitoring, the overheads incurred while running with
% monitoring could be determined. Although this would give an accurate
% measurement of overheads, it would require modifying the program binary. In
% addition, the overhead budget for monitoring would only be updated at
% checkpoints.

% Instead, we develop a model to estimate the execution time overhead budget in a
% gradual manner and without the need to modify the main program. Specifically,
We specify the overhead budget as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles. 
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
If the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events should be dropped.

Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented. This can be difficult to determine precisely due to the
difficulty in measuring certain overheads such as contention for shared memory.
However, we have found that using only the stalls due to FIFO back pressure
works well in practice.

\subsection{Dropping Decision Trade-Offs}
\label{sec:policies.events}

@@

The simplest policy for when to drop monitoring events is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be filtered. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.

@@

An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.
Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. We expect source dropping 
to work well when there are a large number of independent metadata flows.

>>>>

In order to use partial monitoring to enable adjustable overheads, we must also
specify a policy for when and which monitoring events are dropped. In this
section, we discuss some of the options and trade-offs for dropping policies.
We split this decision into two components:
\begin{enumerate}
  \item When do we need to drop events in order to enable reduced overheads?
  \item Which events should be valid for dropping?
\end{enumerate}

Section~\ref{sec:policies.when} discusses options for deciding when events are
dropped while Section~\ref{sec:policies.which} discusses trade-offs for which
events are capable of being dropped.

% Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
% This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
% made. Specifically, we first discuss how to decide when dropping should be performed 
% Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
% selecting which events are dropped (Section~\ref{sec:policies.events}). Different policies in determining which
% events can be dropped create a trade-off between how closely the overhead
% budget is met and the monitoring coverage achieved.

\subsection{Deciding When to Drop}
\label{sec:policies.when}

@@

In this section, we discuss two possible ways to determine when events should be dropped.
The first possibility is to probabilistically drop events. By setting the probability of
dropping events appropriately, overheads can be reduced. This is works well for
enabling partial monitoring for cooperative testing and debugging since the
randomness allows different users and runs to monitor different portions of the
program. However, using a probabilistic dropping policy can make it difficult
to meet a target overhead without prior profiling.

Alternatively, we can specify a target overhead and estimate, at run-time, the
overheads of monitoring in order to decide whether dropping is needed.
% In order to decide whether a monitoring event should be dropped or not, we need
% a way to keep track of overheads. 
We present a specific method
for measuring execution time overheads. However, we note that other methods
could also be used in order to target budgets on IPC, power, energy, etc.  
% For example, one way to measure execution time overheads would be to insert
% checkpoints in the program. By recording the execution time of these
% checkpoints without monitoring, the overheads incurred while running with
% monitoring could be determined. Although this would give an accurate
% measurement of overheads, it would require modifying the program binary. In
% addition, the overhead budget for monitoring would only be updated at
% checkpoints.

% Instead, we develop a model to estimate the execution time overhead budget in a
% gradual manner and without the need to modify the main program. Specifically,
We specify the overhead budget as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles. 
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
In this slack-based policy, if the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events should be dropped.

% % Slack tracking module
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/stm.pdf}
%     \vspace{-0.2in}
%     \caption{Slack tracking and drop decision hardware.}
%     \label{fig:policies.stm}
%     \vspace{-0.1in}
%   \end{center}
% \end{figure}

% Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
% keeping track of slack. 
Slack can be easily measured in hardware by using a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, the measured slack
is decremented. This can be difficult to determine precisely due to the
difficulty in measuring certain overheads such as contention for shared memory.
However, we have found that using only the stalls due to FIFO back pressure
works well in practice.

\subsection{Deciding Which Events to Drop} 
\label{sec:policies.which}

@@

In addition to deciding when dropping is required, trade-offs also exist in
deciding which dropping events can be dropped.
The simplest policy is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be filtered. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.

@@

An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-only dropping} and we refer to the previous policy of 
dropping any event as \emph{unrestricted dropping}.
Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. We expect source dropping 
to work well when there are a large number of independent metadata flows.
In addition, using a probabilistic dropping policy works poorly with
unrestricted dropping since each event in a chain needs to be randomly selected
for monitoring in order for the check to be performed. Instead, probabilistic
dropping works better when coupled with source-only dropping.
